\subsubsection{Blind Assignment for Blockchain Extension (BABE)}
\label{sec:babe}

In Polkadot, we produce relay chain blocks using our Blind Assignment for Blockchain Extension protocol (BABE).
 %\eray{, abbreviated BABE}{(BABE)}.
BABE assigns validators randomly to block production slots using  the randomness generated with blocks. A block production slot is a division of time when a block producer may produce a block. Note, that time is not universally agreed on, which we will address later.  These assignments are completely private until the assigned validators produce their blocks. Therefore, we use ``Blind Assignment'' in the protocol name. BABE is similar to Ouroboros Praos \cite{Praos} with some significant differences in the chain selection rule and timing assumptions.

In BABE, we may have slots without any assignment
%\eray{c}{assinments,}
 which we call empty slot. 
%\eray{slot}{slots}.
In order to fill the empty slots, we have a
%\eray{}{a}
secondary block production mechanism based on Aura \cite{aura} that assigns validators to slots publicly. We note that these blocks do not contribute to
%\eray{}{to} 
the security of BABE since the best chain selection and the random number generation algorithms work as if Aura blocks do not exist.
%\eray{the randomness generation \eray{do}{does} not consider randomness in the Aura blocks.}{-comment:I didn't understand this sentence.-}
Therefore, next we only describe BABE together with its security properties.
%since Aura blocks \eray{is}{are} not the part of the security of BABE.


BABE \cite{babe} consists of another time division called \emph{epochs} ($e_1,e_2,...$), where each epoch consists of a number of sequential block production slots ($e_i = \{sl^i_{1}, sl^i_{2},\ldots,sl^i_{t}\}$) up to the bound  $R$.
Each validator knows in which slots it is supposed to produce a block at the beginning of every epoch. When the time for its slot comes, the validator produces the block by proving that it is assigned to
%\eray{}{to} 
this slot.

The blind assignment is based on the cryptographic primitive called verifiable random function (VRF) \cite{vrf} (see Section \ref{sec:session_keys}). 
A validator in an epoch $e_m$  does the following to learn if it is eligible to produce a block in slot $sl_i^m$:
\begin{enumerate}
	\item  it obtains the randomness in the genesis block if $ m = 1  $ or $ m =2 $. Otherwise, it obtains the randomness  generated two epochs before ($e_{m-2}$).
	\item  it runs the VRF with its secret key and the input:  randomness and the slot number $ sl_i^m $.
\end{enumerate}

If the output of  VRF is less than the threshold $ \tau $, then the validator is the slot leader meaning that it is eligible to produce a block for this slot. We select $\tau$ with respect to security requirements of BABE \cite{babe} e.g., bigger $ \tau $ makes less probable to select only honest validators for a slot than smaller $ \tau $. 
When a validator produces a block, it adds the output of the VRF and its proof to the block which shows that its VRF output is less than $\tau$  in order to convince other validators that it has a right to produce a block in the corresponding slot. The validators always generate their blocks on top of the best chain.
The best chain selection rule in BABE says that ignore the Aura blocks and select the longest chain that includes the last finalised GRANDPA block. See Section \ref{sec:grandpa} for the details how blocks are finalised in GRANDPA.

The randomness of an epoch $e_m$ where $ m > 2 $ is generated by using the BABE blocks of the best chain that belongs to that epoch: let $\rho$ be the concatenation of all  VRF values in BABE blocks that belongs to $e_m$. Then, compute the randomness for epoch $ e_m $ as $r_{m} = H(m
||\rho)$ where $ H $ is a hash function. Validators run periodically the relative time algorithm described below to learn at what time a slot starts according to their local clocks.

\input{relativetime.tex}


\paragraph{Security Overview of BABE:} Garay et al. \cite{backbone} define the properties below in order to obtain a secure blockchain protocol. Informally, we can describe these properties as follows:

\begin{itemize}
	\item \emph{Common Prefix (CP):} \label{item:common_prefix}
	%\eray{}{-comment: Isn't this CPP? How did you get CP as an abbreviation here?-}:}: It is used as CP in the literature. I removed the Property to be consisten with the name of the next properties
	 It ensures that the blocks which are $ k $-blocks before the last block of an honest validator's blockchain cannot be changed. We call  all unchangeable blocks  \emph{finalized} blocks. BABE satisfies CP property thanks to the honest super majority since malicious validators are selected for a slot probabilistically much less than the honest validators. It means that malicious validators do
	 % \eray{does}{do}
	  not have enough 
	  %\eray{source}{sources}: source sounds better for me
	  to construct another chain which does not include one of the finalised blocks.
	\item \emph{Chain Quality (CQ):} \label{item:chain_quality} It ensures a minimum honest block contribution to any best chain owned by an honest party in every certain number of slots. We guarantee even in the worst case where a network delay is maximum that there will be at least one honest block in the best chain during an epoch so that the randomness cannot be biased.
	\item \emph{Chain Growth (CG):} \label{item:chain_growth} It guarantees a minimum growth between slots. Thanks to super majority of honest validators, malicious validators cannot prevent the growth of the best chain.
	
	\item \emph{Chain Density (CD):} \label{item:chain_density} It ensures that in a sufficiently long portion of the best chain more than half of the blocks produced by honest validators. CQ and CG properties %\eray{implies}{imply} 
	imply this property \cite{Praos}.
\end{itemize}
Further details about BABE and  its security analysis can be found in \cite{babe}.

%\eray{Please see \cite{babe} for further details about BABE and  its security analysis.}{Further details about BABE and  its security analysis can be found in \cite{babe}.}
