\subsection{Cryptography}\label{sec:crypto}

In Polkadot, we necessarily distinguish among different permissions and functionalities with different keys and key types, respectively.  We roughly categorise these into account keys with which users interact and session keys that nodes manage without operator intervention beyond a certification process.

\subsubsection{Account keys}

Account keys have an associated balance of which portions can be {\em locked} to play roles in staking, resource rental, and governance, including waiting out a couple types of unlocking period.  We allow several locks of varying duration, both because these roles impose different restrictions, and for multiple unlocking periods running concurrently. 

We encourage active participation in all these roles, but they all require occasional signatures from accounts.  At the same time, account keys have better physical security when kept in inconvenient locations, like safety deposit boxes, which makes signing arduous.  We avoid this friction for users as follows.

Accounts that lock funds for staking are called {\em stash accounts}.  All stash accounts register a certificate on-chain that delegates all validator operation and nomination powers to some {\em controller account}, and also designates some {\em proxy key} for governance votes.  In this state, the controller and proxy accounts can sign for the stash account in staking and governance functions respectively, but not transfer fund.  

\smallskip

At present, we suport both ed25519 \cite{} and schnorrkel/sr25519 \cite{} for account keys.  These are both Schnorr-like signatures implemented using the Ed25519 curve, so both offer extremely similar security.  We recommend ed25519 keys for users who require Hardware Security Module (HSM) support or other external key management solution, while schnorrkel/sr25519 provides more blockchain-friendly functionality like Hierarchical Deterministic Key Derivation (HDKD) and multi-signatures \cite{}.  

In particular, schnorrkel/sr25519 uses the Ristretto implementation \cite{Ristretto} of Mike Hamburg's Decaf \handan{\cite[\S7]{Decaf}}{a strange symbol appears in pdf}, which provide the 2-torsion free points of the Ed25519 curve as a prime order group.  Avoiding the cofactor like this means Ristretto makes implementing more complex protocols significantly safer.  We employ Blake2b for most conventional hashing in Polkadot, but schnorrkel/sr25519 itself uses STROBE128 \cite{STROBE}, which is based on Keccak-f(1600) and provides a hashing interface well suited to signatures and Non-Interactive Zero-Knowledge Proofs (NIZK) \cite{} \cite{}.
% See https://github.com/w3f/schnorrkel/blob/master/annoucement.md for more detailed design notes.

\subsubsection{Session keys}\label{sec:session_keys}

Session keys each fill roughly one particular role in consensus or security.  As a rule, session keys gain authority only from a session certificate, signed by some controller key, that delegates appropriate stake.  

At any time, the controller key can pause or revoke this session certificate and/or issue replacement with new session keys.  All new session keys can be registered in advance, and most must be, so validators can cleanly transition to new hardware by issuing session certificates that only become valid after some future session.  We suggest using pause mechanism for emergency maintenance and using revocation if a session key might be compromised.  

We prefer if session keys remain tied to one physical machine because doing so minimizes the risk of accidental equivocation.  We ask validator operators to issue session certificate using an RPC protocol, not to handle the session secret keys themselves.  

Almost all early proof-of-stake networks have a negligent public key infrastructure that encourages duplicating session secret keys across machines, and thus reduces security and leads to pointless slashing.
% TODO: I'd meant to cite this somewhere, but not sure where now.  It's easy to cite the slashing to part to cosmos.  Thoughts?

\smallskip

We impose no prior restrictions on the cryptography employed by specific components or their associated session keys types.\footnote{We always implement cryptography for polkadot in native code, not just because the runtime suffers from WASM's performance penalties, but because all of Polkadot's consensus protocols are partially implemented outside the runtime in Substrate modules.}

In BABE \ref{sec:babe}, validators use schnorrkel/sr25519 keys both for regular Schnorr signatures, as well as for a verifiable random function (VRF) based on NSEC5 \cite{NSEC5}.  

A VRF is the public-key analog of a pseudo-random function (PRF), aka cryptographic hash function with a distinguished key, such as many MACs.  We award block production slots when the block producer scores a low enough VRF output $\mathtt{VRF}_{\sk}(r_e || \mathtt{slot_number} )$, so anyone with the VRF public keys can verify that blocks were produced in the correct slot, but only the block producers know their slots in advance via their VRF secret key.

As in \cite{Praos}, we provide a source of randomness $r_e$ for the VRF inputs by hashing together all VRF outputs form the previous session, which requires that BABE keys be registered at least two full epochs before being used.

We reduce VRF output malleability by hashing the signer's public key along side the input, which dramatically improves security when used with HDKD.  We also hash the VRF input and output together when providing output used elsewhere, which improves compossibility when used as a random oracle in security proofs.  See the 2Hash-DH construction from Theorem 2 on page 32 in appendix C of \cite{Praos}.  

In GRANDPA \ref{sec:grandpa}, validators shall vote using BLS signatures, which supports convenient signature aggregation and select ZCash's BLS12-381 curve for performance.  There is a risk that BLS12-381 might drops significantly below 128 bits of security, due to number field sieve advancements.  If and when this happens, we expect upgrading GRANDPA to another curve to be straightforward. 
% https://mailarchive.ietf.org/arch/msg/cfrg/eAn3_8XpcG4R2VFhDtE_pomPo2Q

% TODO: ImOnline
% ref. https://github.com/paritytech/substrate/issues/3546

We treat libp2p's transport keys roughly like session keys too, but they include the transport keys for sentry nodes, not just for the validator itself.  As such, the operator interacts slightly more with these.

We permit controller keys to revoke session key validity of course, but controllers could pause operation for shorter periods.  We similarly permit controllers to register new session keys in advance, which enables a clean handover between validator machines.

