\section{Preliminaries}

\subsection{Adversarial Model of Polkadot}




The adversarial assumption on collators depends on the parachain type. If the parachain is private, we assume that there is at least one honest collator. Otherwise, we assume that at least the majority of collators are honest.

\paragraph{Keys:} We assume that malicious parties generate their keys with an arbitrary algorithm while honest ones always generate their keys securely.

\paragraph{Network and Communication:} All validators have their local clock. We do not rely on any central clock in Polkadot. We assume that the network of validators and collators is partially synchronous. It means that a message sent by a validator or collator arrives at all parties in the network at most $\D$ times later where $\D$ is an unknown parameter. So, we assume an eventual delivery in Polkadot.
We also assume that collators and fishermen can connect to the relay chain network to submit their reports.

\subsection{Roles}
A fixed set of actors called \emph{Validators} are responsible to maintain the relay chain.
These validators need to vote on the consensus over all the parachains.
In addition, the validators are assigned to parachain by dividing the validators set into
random disjoint subsets that are each assigned to a particular parachain.
For the parachain, there are additional actors called collators and fishermen that are
responsible for parachain block production and reporting invalid parachain blocks respectively.