
\section{Others}

{\bf ALL TEXT HERE LIFTED FROM PREVIOUS VERSION, SOME OF IT IS WRONG.}


\subsection{Validators}

\emph{Slashing:} If at least $f+1$ validators sign for the invalidity of a block and less than $f+1$ validators sign for the validity of the block, then we slash all validators who signed for the validity.

If at least $f+1$ validators sign for the validity of a block and less than $f+1$ validator sign for invalidity, we slash the ones who signs for the invalidity.

We note that $f+1$ validity signatures and $f+1$ invalidity signatures are not possible given that at most $f$ validators are malicious and proof of block algorithms  are correct (See Definition \ref{def:pob}).

\emph{Rewarding:} All parachain validators and extra validators who signed for the final decision (either valid or invalid) of a block receive a reward. 

Eligibility of the reward can be easily checked for the parachain validators and extra validators who satisfies the conditions when $\tau = 0$. Latter, we just need to verify the VRF proof.
However, it is not easy to verify the condition for $\tau > 0$, because this value is subjective. Therefore, we may not need to be very precise for this $\tau$ value, if the signature of the validator is correct.


\subsection{Fishermen}
\label{sec:fishermen}

TODO: Comment on LDPC codes from Coded Merkle Trees paper

...

All fishermen collect blobs from collators in order to check the validity all the time. Whenever they discover an invalid block, they announce it to the validators by signing their claim. These claims are added to the block list by the validators. So, these claims are going to be in the relay chain. They do not provide any proof. If their claims are wrong then they are slashed. Otherwise, they are rewarded. The correctness of the fishermen claims is determined with the validity check protocol.

...

\emph{Slashing:} If at least $\totalchecks$ validators sign saying that the block is valid, then fishermen with the claim saying the block is invalid are slashed. All the claims from these fishermen are ignored in future.

\emph{Rewarding:} If $f+1$ validators sign to say the block is invalid, the  fishermen with the claim saying that the block is invalid are rewarded.

\subsection{Collators}
\label{sec:collators}

We expect collators are free to act as fishermen ..
except collator authority should probably be additive with fishermen authority 

... unavailability reports from collators ...

